---
title: "TayChenJiangNguyenLi-stat443project"
author: "TayChenJiangNguyenLi"
date: "2026-03-16"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages
```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(lubridate)
```

## Reading the data
```{r, results='hide'}
# Covering 1981-01 to 2025-12
nhpi_full <- read.csv("18100205.csv", na.strings = c("..", "NA", "")) %>%
  mutate(REF_DATE = as.Date(paste0(REF_DATE, "-01")),
         REF_DATE_ym = format(REF_DATE, "%Y-%m"))
# head(nhpi_full)

# Covering 1981-01 to 2023-12
ir <- read_csv("IRSTCB01CAM156N.csv",show_col_types = FALSE) %>%
  rename(REF_DATE = observation_date,
         IR = IRSTCB01CAM156N) %>%
  mutate(REF_DATE = as.Date(REF_DATE),
         REF_DATE_ym = format(REF_DATE, "%Y-%m"),
         IR = as.numeric(IR)) %>%
  filter(!is.na(IR))
# head(ir)

# Combining the dataset for ARMAX (cover the overlap only)
nhpi_ir <- nhpi_full %>%
  left_join(ir %>% select(REF_DATE_ym, IR),
            by = "REF_DATE_ym") %>%
  filter(REF_DATE <= as.Date("2023-12-01"))
# head(nhpi_ir)

# Getting lagged interest rate
nhpi_ir <- nhpi_ir %>%
  group_by(GEO, New.housing.price.indexes) %>%
  arrange(REF_DATE) %>%
  mutate(IR_lag1 = lag(IR, 1),
         IR_lag2 = lag(IR, 2)) %>%
  ungroup()
# head(nhpi_ir)

# Checking date coverage
range(nhpi_full$REF_DATE)  # should include 1981-01-01 to 2025-12-01 (or close)
range(ir$REF_DATE)         # should include 1981-01-01 to 2023-12-01
range(nhpi_ir$REF_DATE)    # should include 1981-01-01 to 2023-12-01

# Checking if there's any missing IR after merge? (should be basically none within overlap)
mean(is.na(nhpi_ir$IR))

# Checking lags (first month per series will be NA)
nhpi_ir %>%
  filter(GEO == "Canada", New.housing.price.indexes == "Total (house and land)") %>%
  select(REF_DATE, VALUE, IR, IR_lag1, IR_lag2) # %>%
  # head(5)
```

## Defining functions

### Baseline 1: Naive/Persistence
```{r}
persist_fc <- function(train, holdout, iprint = FALSE) {
  n <- length(train)
  n_holdout <- length(holdout)
  mse <- 0
  fcvec <- numeric(n_holdout)
  
  # first forecast
  fc <- train[n]
  yt <- holdout[1]
  fcvec[1] <- fc
  fcerror <- yt - fc
  mse <- mse + fcerror^2
  
  # remaining forecasts
  for (i in 2:n_holdout) {
    yt <- holdout[i]
    fc <- holdout[i-1]
    fcvec[i] <- fc
    fcerror <- yt - fc
    mse <- mse + fcerror^2
  }
  
  if (iprint) 
    print(cbind(holdout, fcvec))
  
  rmse <- sqrt(mse / n_holdout)
  return(list(fc = fcvec, rmse = rmse))
}
```

### Baseline 2: Training-Set Average
```{r}
iid_fc <- function(train, holdout, iprint = FALSE) {
  n <- length(train)
  n_holdout <- length(holdout)
  average <- mean(train)
  mse <- 0
  fcvec <- numeric(n_holdout)

  for (i in 1:n_holdout) {
    yt <- holdout[i]
    fc <- average
    fcvec[i] <- fc
    fcerror <- yt - fc
    mse <- mse + fcerror^2
  }
  
  if (iprint)
    print(cbind(holdout, fcvec))
  
  rmse <- sqrt(mse / n_holdout)
  return(list(fc = fcvec, rmse = rmse))
}
```

### Method 1: Best Exponential Smoothing (allowing for trend and seasonality, if relevant)
```{r}
# Simple Exponential Smoothing
esm_fc <- function(train, holdout, alpha, level, iprint = FALSE) {
  n_holdout <- length(holdout)
  sse <- 0
  fcvec <- numeric(n_holdout)
  
  fc <- level
  yt <- holdout[1]
  level_new <- level
  fcvec[1] <- fc
  fcerror <- yt - fc
  sse <- sse + fcerror^2
  
  for (i in 2:n_holdout) {
    level_new <- alpha * holdout[i-1] + (1 - alpha) * level_new
    
    fc <- level_new
    fcvec[i] <- fc
    yt <- holdout[i]
    fcerror <- yt - fc
    sse <- sse + fcerror^2
    
    if (iprint) {
      cat("i =", i, "| fc =", fc, "| y =", yt, "| level =", level_new, "\n")
    }
  }
  
  rmse <- sqrt(sse / n_holdout)
  return(list(fc = fcvec, rmse = rmse))
}
```

```{r}
# Linear Exponential Smoothing
lholt_fc <- function(train, holdout, alpha, beta, level, slope, iprint = FALSE) {
  n_holdout <- length(holdout)
  sse <- 0
  fcvec <- numeric(n_holdout)
  
  # first forecast
  fc <- level + slope
  yt <- holdout[1]
  newfc <- fc
  fcvec[1] <- fc
  fcerror <- yt - fc
  sse <- sse + fcerror^2
  
  level_prev <- level
  slope_prev <- slope
  
  if (iprint) {
    cat("i = 1 | fc =", fc, "| y =", yt, "| level =", level_prev, "| slope =", slope_prev, "\n")
  }
  
  # remaining forecasts
  for (i in 2:n_holdout) {
    level_new <- alpha * holdout[i-1] + (1 - alpha) * fc
    slope_new <- beta * (level_new - level_prev) + (1 - beta) * slope_prev
    
    fc <- level_new + slope_new
    fcvec[i] <- fc
    yt <- holdout[i]
    fcerror <- yt - fc
    sse <- sse + fcerror^2
    
    if (iprint) {
      cat("i =", i, "| fc =", fc, "| y =", yt, "| level =", level_new, "| slope =", slope_new, "\n")
    }
    
    level_prev <- level_new
    slope_prev <- slope_new
  }
  
  rmse <- sqrt(sse / n_holdout)
  return(list(fc = fcvec, rmse = rmse))
}
```

```{r}
# Additive Seasonal Exponential Smoothing
aseason_fc <- function(train, holdout, alpha, beta, gamma, level, slope, season, iprint = FALSE) {
  n_holdout <- length(holdout)
  m <- length(season)
  sse <- 0
  fcvec <- numeric(n_holdout)
  
  # prev states from training end
  level_prev <- level
  slope_prev <- slope
  
  # helper: map i -> seasonal position in 1..d
  spos <- function(i) ((i-1) %% m) + 1
  
  # First forecast
  fc <- level_prev + slope_prev + season[spos(1)]
  fcvec[1] <- fc
  yt <- holdout[1]
  fcerror <- yt - fc
  sse <- sse + fcerror^2
  
  if (iprint) {
    cat("i = 1 | fc =", fc, "| y =", yt, "| level =", level_prev, "| slope =", slope_prev, "| season =", season[spos(1)], "\n")
  }
  
  # Remaining forecast
  for (i in 2:n_holdout) {
    # Seasonal of previous observation (i-1)
    k_prev <- spos(i-1)
    s_prev <- season[k_prev]
    # Update using previous observation holdout[i-1]
    level_new <- alpha * (holdout[i-1] - s_prev) + (1 - alpha) * (level_prev + slope_prev)
    slope_new <- beta * (level_new - level_prev) + (1 - beta) * slope_prev
    s_new <- gamma * (holdout[i-1] - level_new) + (1 - gamma) * s_prev
    
    # store updated seasonal state one full cycle later
    season[k_prev] <- s_new
    
    # Seasonal of current forecast (i)
    k_cur <- spos(i)
    s_cur <- season[k_cur]
    
    # Forecast current observation using updated states + current seasonal
    fc <- level_new + slope_new + s_cur
    fcvec[i] <- fc
    yt <- holdout[i]
    fcerror <- yt - fc
    sse <- sse + fcerror^2
    
    if (iprint) {
      cat("i =", i, "| fc =", fc, "| y =", yt, "| level =", level_new, "| slope =", slope_new, "| season =", s_cur, "\n")
    }
    
    # Update level and slope
    level_prev <- level_new
    slope_prev <- slope_new
  }
  
  rmse <- sqrt(sse / n_holdout)
  return(list(fc = fcvec, rmse = rmse))
}
```

### Method 2: ARIMA
```{r}
arma_fc <- function(tsdata, ntrain, order, seasonal, method, traincoef, include.mean, iprint=F) {
  obj = arima(tsdata, order=order, seasonal = seasonal, init=traincoef, fixed=traincoef,
              method=method, include.mean=include.mean, optim.control=list(maxit=0)) # no log-likelihood iterations
  fc = tsdata - obj$residuals

  ntotal = length(tsdata)
  holdout_fc = fc[(ntrain+1):ntotal]
  holdout = tsdata[(ntrain+1):ntotal]
  
  if(iprint) print(cbind(holdout,holdout_fc))
  
  rmse = sqrt(mean((holdout-holdout_fc)^2))
  return(list(rmse=rmse, fc=holdout_fc, residuals = obj$residuals))
}
```

### Method 3: ARIMAX with lagged interest
```{r}
armax_fc <- function(tsdata, ntrain, order, seasonal, xreg, method, traincoef, include.mean, iprint=F) { 
  obj = arima(tsdata, order=order, seasonal = seasonal, init=traincoef,
              fixed=traincoef, method=method, include.mean=include.mean, 
              xreg=xreg, optim.control=list(maxit=0))

  fc = tsdata - obj$residuals
  
  ntotal = length(tsdata)
  holdout_fc = fc[(ntrain+1):ntotal]
  holdout = tsdata[(ntrain+1):ntotal]
  
  if(iprint) print(cbind(holdout,holdout_fc))
  
  rmse = sqrt(mean((holdout-holdout_fc)^2))
  return(list(rmse=rmse, fc=holdout_fc, residuals = obj$residuals))
}
```

## Analyzing Montreal

### Setting up
```{r}
# Reading the data
montreal_nhpi_data <- read.csv("Full_Cities/montreal_full.csv")
montreal_nhpi_ir_data <- read.csv("Extracted_Cities/montreal.csv")

# Extracting the NHPI
montreal_nhpi <- montreal_nhpi_data$VALUE

# Extracting the one-month lag interest rate
montreal_irlag1 <- montreal_nhpi_ir_data$IR_lag1

# Setting the time window from 1981-2023
montreal_ts <- ts(montreal_nhpi, start = c(1981,1), end = c(2023, 12), frequency = 12)
montreal_ir_ts <- ts(montreal_irlag1, start = c(1981,1), end = c(2023, 12), frequency = 12)

# Applying a transformation for NHPI
montreal_transformed_nhpi <- diff(log(montreal_ts))

# Setting a training set from 1981-2021
montreal_train_ts <- window(montreal_transformed_nhpi, end = c(2021,12))
montreal_train_ir_xreg <- window(montreal_ir_ts, start = c(1981,2), end = c(2021,12))

# Setting a holdout set from 2022-2023
montreal_holdout_ts <- window(montreal_transformed_nhpi, start = c(2022,1))
montreal_holdout_ir_xreg <- window(montreal_ir_ts, start = c(2022,1))
```

### Doing the baseline forecasts
```{r}
# Baseline 1: Naive/Persistence
montreal_persist_fc <- persist_fc(montreal_train_ts, montreal_holdout_ts, iprint = F)

# Baseline 2: Training-Set Average
montreal_iid_fc <- iid_fc(montreal_train_ts, montreal_holdout_ts, iprint = F)
```

### Doing the exponential smoothing forecasts
```{r}
# Method 1
ntrain <- length(montreal_train_ts)
ntotal <- length(montreal_ts)-1

montreal_transformed_nhpi_train <- montreal_transformed_nhpi[1:ntrain]
montreal_transformed_nhpi_holdout <- montreal_transformed_nhpi[(ntrain+1):ntotal]

# Simple Exponential Smoothing
fit_expsmo <- HoltWinters(montreal_train_ts, beta = F, gamma = F)
montreal_esm_fc <- esm_fc(montreal_transformed_nhpi_train, montreal_transformed_nhpi_holdout,
                         alpha = fit_expsmo$alpha, level = fit_expsmo$coefficients["a"],
                         iprint = F)

# Linear Exponential Smoothing
fit_hw <- HoltWinters(montreal_train_ts, gamma = F)
montreal_holt_fc <- lholt_fc(montreal_transformed_nhpi_train, montreal_transformed_nhpi_holdout,
                           alpha = fit_hw$alpha, beta = fit_hw$beta, 
                           level = fit_hw$coefficients["a"], 
                           slope = fit_hw$coefficients["b"], iprint = F)

# Additive Seasonal Exponential Smoothing
wafit <- HoltWinters(montreal_train_ts, seasonal = "additive")
montreal_aseason_fc <- aseason_fc(montreal_transformed_nhpi_train,
                                 montreal_transformed_nhpi_holdout, 
                                 alpha = wafit$alpha, beta = wafit$beta, 
                                 gamma = wafit$gamma, level = wafit$coef[1], 
                                 slope = wafit$coef[2], season = wafit$coefficient[3:14],
                                 iprint = F)

# Comparing RMSE
rmse_esm <- cbind(montreal_esm_fc$rmse, montreal_holt_fc$rmse, montreal_aseason_fc$rmse)
colnames(rmse_esm) <- c("esm", "lholt", "aseason") # RMSE: 0.0079, 0.0077, 0.0088
rmse_esm

# Best exponential smoothing method: Linear Exponential Smoothing
```

### Doing the ARIMA forecasts
```{r, warnings=FALSE}
# Method 2
montreal_transformed_nphi_auto <- forecast::auto.arima(montreal_train_ts)
summary(montreal_transformed_nphi_auto) # 0.0048783
resid_fit <- residuals(montreal_transformed_nphi_auto)
acf(resid_fit, main = "ACF of ARIMA(2,0,5)(2,0,1)[12] residuals")

# Comparing with the other ARIMA models
manual_fit <- arima(montreal_train_ts, order = c(2,0,5),
                    seasonal = list(order = c(2,0,1), period = 12), method = "ML")
summary(manual_fit) # RMSE: 0.004903117
acf(residuals(manual_fit), main = "ACF of ARIMA(2,0,5)(2,0,1)[12] residuals")
test1 <- arima(montreal_train_ts, order = c(2,0,4),
               seasonal = list(order = c(2,0,1), period = 12), method = "ML")
summary(test1) # RMSE: 0.004919859
acf(residuals(test1), main = "ACF of ARIMA(2,0,4)(2,0,1)[12] residuals")
test2 <- arima(montreal_train_ts, order = c(3,0,4),
               seasonal = list(order = c(2,0,1), period = 12), method = "ML")
summary(test2) # RMSE: 0.004857711
acf(residuals(test2), main = "ACF of ARIMA(3,0,4)(2,0,1)[12] residuals")
test3 <- arima(montreal_train_ts, order = c(3,0,5),
               seasonal = list(order = c(2,0,1), period = 12), method = "ML")
summary(test3) # RMSE: 0.004852676
acf(residuals(test3), main = "ACF of ARIMA(3,0,5)(2,0,1)[12] residuals")
test4 <- arima(montreal_train_ts, order = c(2,0,5),
               seasonal = list(order = c(2,0,2), period = 12), method = "ML")
summary(test4) # RMSE: 0.004908301
acf(residuals(test4), main = "ACF of ARIMA(2,0,5)(2,0,2)[12] residuals")
test5 <- arima(montreal_train_ts, order = c(2,0,4),
               seasonal = list(order = c(2,0,2), period = 12), method = "ML")
summary(test5) # RMSE: 0.004891118
acf(residuals(test5), main = "ACF of ARIMA(2,0,4)(2,0,2)[12] residuals")

# Comparing AIC values
aic = cbind(manual_fit$aic, test1$aic, test2$aic, test3$aic, test4$aic, test5$aic)
colnames(aic) = c("manual_fit", "test1", "test2", "test3", "test4", "test5")
aic
# Best (based on lowest AIC and ACF plot): test3, ARIMA(3,0,5)(2,0,1)[12]

order <- c(3,0,5)
seasonal <- list(order = c(2,0,1), period = 12)
# Fit the model to the training data
train_arima_model <- arima(montreal_train_ts, order = order,
                           seasonal = seasonal, method = "ML", include.mean = T)
# Result the moving 1-step ahead forecast for holdout set
arima_results <- arma_fc(tsdata = montreal_transformed_nhpi,
                         ntrain = length(montreal_train_ts),
                         order = order, seasonal = seasonal, method = "ML",
                         traincoef = coef(train_arima_model), include.mean = T)
arima_results$rmse
```
### Compare candidate based on holdout rmse
```{r}
# Compare candidate ARIMA models using holdout RMSE
rmse_manual_fit <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(2,0,5),
  seasonal = list(order = c(2,0,1), period = 12),
  method = "ML",
  traincoef = coef(manual_fit),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test1 <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(2,0,4),
  seasonal = list(order = c(2,0,1), period = 12),
  method = "ML",
  traincoef = coef(test1),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test2 <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(3,0,4),
  seasonal = list(order = c(2,0,1), period = 12),
  method = "ML",
  traincoef = coef(test2),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test3 <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(3,0,5),
  seasonal = list(order = c(2,0,1), period = 12),
  method = "ML",
  traincoef = coef(test3),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test4 <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(2,0,5),
  seasonal = list(order = c(2,0,2), period = 12),
  method = "ML",
  traincoef = coef(test4),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test5 <- arma_fc(
  tsdata = montreal_transformed_nhpi,
  ntrain = length(montreal_train_ts),
  order = c(2,0,4),
  seasonal = list(order = c(2,0,2), period = 12),
  method = "ML",
  traincoef = coef(test5),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

# Put all candidate RMSE values into one table
arima_rmse_compare <- data.frame(
  model = c(
    "ARIMA(2,0,5)(2,0,1)[12]",
    "ARIMA(2,0,4)(2,0,1)[12]",
    "ARIMA(3,0,4)(2,0,1)[12]",
    "ARIMA(3,0,5)(2,0,1)[12]",
    "ARIMA(2,0,5)(2,0,2)[12]",
    "ARIMA(2,0,4)(2,0,2)[12]"
  ),
  rmse = c(
    rmse_manual_fit,
    rmse_test1,
    rmse_test2,
    rmse_test3,
    rmse_test4,
    rmse_test5
  )
)

arima_rmse_compare

best_arima_model_M <- "ARIMA(2,0,5)(2,0,2)[12]"
best_arima_rmse_M <- 0.007296657

```


### Doing the ARIMAX forecasts with lagged interest
```{r}
# Method 3
montreal_fit_arimax <- forecast::auto.arima(montreal_train_ts, xreg = montreal_train_ir_xreg)
summary(montreal_fit_arimax)

order <- c(2,0,5)
seasonal <- list(order = c(2,0,2), period = 12)
# Fitting a model to the training data
train_arimax_model <- arima(montreal_train_ts, order = order, seasonal = seasonal,
                            xreg = montreal_train_ir_xreg, method = "ML", include.mean = T)
# Getting the result of the moving 1-step ahead forecast for holdout set
arimax_results <- armax_fc(tsdata = montreal_transformed_nhpi,
                           ntrain = length(montreal_train_ts), order = order,
                           seasonal = seasonal, xreg = montreal_irlag1[-1],
                           method = "ML", traincoef = coef(train_arimax_model), include.mean = T)
montreal_arima_results <- arima_results
montreal_arimax_results <- arimax_results

best_arimax_model_M <- "ARIMAX(2,0,5)(2,0,2)[12]" 
best_arimax_rmse_M  <- montreal_arimax_results$rmse

montreal_best <- data.frame(
  City = "Montreal",
  ARIMA_Model = best_arima_model_M,
  ARIMA_RMSE = round(best_arima_rmse_M, 3),
  ARIMAX_Model = best_arimax_model_M,
  ARIMAX_RMSE = round(best_arimax_rmse_M, 3)
)
```
### Viewing the results of Montreal's analyses
```{r}
# 1. TIME SERIES PLOT
plot(montreal_ts)
# log-diff transformation
# plot(montreal_transformed_ts) # NOTE: this object doesn't exist
# ts plot from persistence
plot(ts(montreal_persist_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from average
plot(ts(montreal_iid_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from simple exponential smoothing
plot(ts(montreal_esm_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from arima
plot(ts(arima_results$fc, start = c(2022, 1), frequency = 12))
# ts plot from arimax
plot(ts(arimax_results$fc, start = c(2022, 1), frequency = 12))
# ts plot of holdout set
plot(ts(montreal_holdout_ts, start = c(2022, 1), frequency = 12))

# 2. ACF PLOT
# ACF and PACF of training set
acf(montreal_train_ts, main = "ACF Plot: of Log-differenced NHPI: Montreal Training Set")
pacf(montreal_train_ts, main = "PACF Plot: of Log-differenced NHPI: Montreal Training Set")
# ACF from best ARIMA model
resid_fit <- residuals(montreal_transformed_nphi_auto)
acf(resid_fit, main = "ACF of ARIMA(2,0,5)(2,0,2)[12] residuals")
# ACF from best ARIMAX model
resid_fit <- residuals(montreal_fit_arimax)
acf(resid_fit, main = "ACF of ARIMAX(2,0,5)(2,0,2)[12] residuals")

# 3. SUMMARY STATISTICS
summary(montreal_train_ts)
summary(montreal_holdout_ts)
# summary of best ARIMA model
summary(montreal_transformed_nphi_auto)
# summary of best ARIMAX model
summary(montreal_fit_arimax)
# CCF between log-differenced NHPI and the lagged interest rate
ccf(montreal_irlag1[-1], diff(log(montreal_nhpi)))

# 4. RMSE
rmse = cbind(montreal_persist_fc$rmse, montreal_iid_fc$rmse,
             montreal_esm_fc$rmse, arima_results$rmse, arimax_results$rmse)
colnames(rmse) = c("persist", "iid", "esm", "arima", "arimax")
print(round(rmse, 3))
```

## Analyzing Toronto

### Setting up
```{r}
# Reading the data
toronto_nhpi_data <- read.csv("Full_Cities/toronto_full.csv")
toronto_nhpi_ir_data <- read.csv("Extracted_Cities/toronto.csv")

# Extracting the NHPI
toronto_nhpi <- toronto_nhpi_data$VALUE

# Extracting the one-month lag interest rate
toronto_irlag1 <- toronto_nhpi_ir_data$IR_lag1

# Setting the time window from 1981-2023
toronto_ts <- ts(toronto_nhpi, start = c(1981,1), end = c(2023, 12), frequency = 12)
toronto_ir_ts <- ts(toronto_irlag1, start = c(1981,1), end = c(2023, 12), frequency = 12)

# Applying a transformation for NHPI
toronto_transformed_nhpi <- diff(log(toronto_ts))

# Setting a training set from 1981-2021
toronto_train_ts <- window(toronto_transformed_nhpi, end = c(2021,12))
toronto_train_ir_xreg <- window(toronto_ir_ts, start = c(1981,2), end = c(2021,12))

# Setting a holdout set from 2022-2023
toronto_holdout_ts <- window(toronto_transformed_nhpi, start = c(2022,1))
toronto_holdout_ir_xreg <- window(toronto_ir_ts, start = c(2022,1))
```

### Doing the baseline forecasts
```{r}
# Baseline 1 
toronto_persist_fc <- persist_fc(toronto_train_ts, toronto_holdout_ts, iprint = F)

# Baseline 2
toronto_iid_fc <- iid_fc(toronto_train_ts, toronto_holdout_ts, iprint = F)
```

### Doing the exponential smoothing forecasts
```{r}
# Method 1
ntrain <- length(toronto_train_ts)
ntotal <- length(toronto_ts)-1

toronto_transformed_nhpi_train <- toronto_transformed_nhpi[1:ntrain]
toronto_transformed_nhpi_holdout <- toronto_transformed_nhpi[(ntrain+1):ntotal]

# Simple Exponential Smoothing
fit_expsmo <- HoltWinters(toronto_train_ts, beta = F, gamma = F)
toronto_esm_fc <- esm_fc(toronto_transformed_nhpi_train, toronto_transformed_nhpi_holdout,
                         alpha = fit_expsmo$alpha, level = fit_expsmo$coefficients["a"],
                         iprint = F)

# Linear Exponential Smoothing
fit_hw <- HoltWinters(toronto_train_ts, gamma = F)
toronto_holt_fc <- lholt_fc(toronto_transformed_nhpi_train, toronto_transformed_nhpi_holdout,
                           alpha = fit_hw$alpha, beta = fit_hw$beta, 
                           level = fit_hw$coefficients["a"], 
                           slope = fit_hw$coefficients["b"], iprint = F)

# Additive Seasonal Exponential Smoothing
wafit <- HoltWinters(toronto_train_ts, seasonal = "additive")
toronto_aseason_fc <- aseason_fc(toronto_transformed_nhpi_train,
                                 toronto_transformed_nhpi_holdout, 
                                 alpha = wafit$alpha, beta = wafit$beta, 
                                 gamma = wafit$gamma, level = wafit$coef[1], 
                                 slope = wafit$coef[2], season = wafit$coefficient[3:14],
                                 iprint = F)

# Comparing RMSE
rmse_esm <- cbind(toronto_esm_fc$rmse, toronto_holt_fc$rmse, toronto_aseason_fc$rmse)
colnames(rmse_esm) <- c("esm", "lholt", "aseason")
rmse_esm

# Best exponential smoothing method: Simple Exponential Smoothing
```

### Doing the ARIMA forecasts
```{r}
# Method 2
toronto_transformed_nphi_auto <- forecast::auto.arima(toronto_train_ts)
summary(toronto_transformed_nphi_auto) # 0.0064057
resid_fit <- residuals(toronto_transformed_nphi_auto)
acf(resid_fit, main = "ACF of ARIMA(1,0,2)(1,0,0)[12] residuals")

# Comparing with the other ARIMA models
manual_fit <- arima(toronto_train_ts, order = c(1,0,2),
                    seasonal = list(order = c(1,0,0), period = 12), method = "ML")
summary(manual_fit) # RMSE: 0.006405747
acf(residuals(manual_fit), main = "ACF of ARIMA(1,0,2)(1,0,0)[12] residuals")
test1 <- arima(toronto_train_ts, order = c(1,0,1),
               seasonal = list(order = c(1,0,0), period = 12), method = "ML")
summary(test1) # RMSE: 0.006456026
acf(residuals(test1), main = "ACF of ARIMA(1,0,1)(1,0,0)[12] residuals")
test2 <- arima(toronto_train_ts, order = c(2,0,1),
               seasonal = list(order = c(1,0,0), period = 12), method = "ML")
summary(test2) # RMSE: 0.006410017
acf(residuals(test2), main = "ACF of ARIMA(2,0,1)(1,0,0)[12] residuals")
test3 <- arima(toronto_train_ts, order = c(2,0,2),
               seasonal = list(order = c(1,0,0), period = 12), method = "ML")
summary(test3) # RMSE: 0.006405754
acf(residuals(test3), main = "ACF of ARIMA(2,0,2)(1,0,0)[12] residuals")
test4 <- arima(toronto_train_ts, order = c(1,0,2),
               seasonal = list(order = c(1,0,1), period = 12), method = "ML")
summary(test4) # RMSE: 0.006404998
acf(residuals(test4), main = "ACF of ARIMA(1,0,2)(1,0,1)[12] residuals")
test5 <- arima(toronto_train_ts, order = c(1,0,1),
               seasonal = list(order = c(1,0,1), period = 12), method = "ML")
summary(test5) # RMSE: 0.006454211
acf(residuals(test5), main = "ACF of ARIMA(1,0,1)(1,0,1)[12] residuals")

# Comparing AIC values
aic = cbind(manual_fit$aic, test1$aic, test2$aic, test3$aic, test4$aic, test5$aic)
colnames(aic) = c("manual_fit", "test1", "test2", "test3", "test4", "test5")
aic
# Best (based on lowest AIC and ACF plot): manual_fit, ARIMA(1,0,2)(1,0,0)[12]

order <- c(1,0,2)
seasonal <- list(order = c(1,0,0), period = 12)
# Fitting the model to the training data
train_arima_model <- arima(toronto_train_ts, order = order,
                           seasonal = seasonal, method = "ML", include.mean = T)
# Getting the result of the moving 1-step ahead forecast for holdout set
arima_results <- arma_fc(tsdata = toronto_transformed_nhpi,
                         ntrain = length(toronto_train_ts), order = order,
                         seasonal = seasonal, method = "ML",
                         traincoef = coef(train_arima_model), include.mean = T)
arima_results$rmse
```
### Compare models based on holdout rmse
```{r}
# Compare candidate ARIMA models using holdout RMSE

rmse_manual_fit <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(manual_fit),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test1 <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(1,0,1),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(test1),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test2 <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(2,0,1),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(test2),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test3 <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(test3),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test4 <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(1,0,2),
  seasonal = list(order = c(1,0,1), period = 12),
  method = "ML",
  traincoef = coef(test4),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_test5 <- arma_fc(
  tsdata = toronto_transformed_nhpi,
  ntrain = length(toronto_train_ts),
  order = c(1,0,1),
  seasonal = list(order = c(1,0,1), period = 12),
  method = "ML",
  traincoef = coef(test5),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

# Put all candidate RMSE values into one table
arima_rmse_compare <- data.frame(
  model = c(
    "ARIMA(1,0,2)(1,0,0)[12]",
    "ARIMA(1,0,1)(1,0,0)[12]",
    "ARIMA(2,0,1)(1,0,0)[12]",
    "ARIMA(2,0,2)(1,0,0)[12]",
    "ARIMA(1,0,2)(1,0,1)[12]",
    "ARIMA(1,0,1)(1,0,1)[12]"
  ),
  rmse = c(
    rmse_manual_fit,
    rmse_test1,
    rmse_test2,
    rmse_test3,
    rmse_test4,
    rmse_test5
  )
)

arima_rmse_compare

best_arima_model_T <- "ARIMA(1,0,2)(1,0,0)[12]"
best_arima_rmse_T <- 0.002048750

```

### Doing the ARIMAX forecasts with lagged interest
```{r}
# Method 3
toronto_fit_arimax <- forecast::auto.arima(toronto_train_ts, xreg = toronto_train_ir_xreg)
summary(toronto_fit_arimax)

order <- c(1,0,2)
seasonal <- list(order = c(1,0,0), period = 12)
# Fitting a model to the training data
train_arimax_model <- arima(toronto_train_ts, order = order, seasonal = seasonal, xreg = toronto_train_ir_xreg, method = "ML", include.mean = T)
# Getting the result of the moving 1-step ahead forecast for holdout set
arimax_results <- armax_fc(tsdata = toronto_transformed_nhpi, ntrain = length(toronto_train_ts), order = order, seasonal = seasonal, xreg = toronto_irlag1[-1], method = "ML", traincoef = coef(train_arimax_model), include.mean = T)
toronto_arima_results <- arima_results
toronto_arimax_results <- arimax_results

best_arimax_model_T <- "ARIMAX(1,0,2)(1,0,0)[12]" 
best_arimax_rmse_T  <- toronto_arimax_results$rmse

toronto_best <- data.frame(
  City = "Toronto",
  ARIMA_Model = best_arima_model_T,
  ARIMA_RMSE = round(best_arima_rmse_T, 3),
  ARIMAX_Model = best_arimax_model_T,
  ARIMAX_RMSE = round(best_arimax_rmse_T, 3)
)
```

### Viewing the results of Toronto's analyses
```{r}
# 1. TIME SERIES PLOT
plot(toronto_ts)
# log-diff transformation
# plot(toronto_transformed_ts) # NOTE: this object doesn't exist
# ts plot from persistence
plot(ts(toronto_persist_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from average
plot(ts(toronto_iid_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from simple exponential smoothing
plot(ts(toronto_esm_fc$fc, start = c(2022, 1), frequency = 12))
# ts plot from ARIMA
plot(ts(arima_results$fc, start = c(2022, 1), frequency = 12))
# ts plot from ARIMAX
plot(ts(arimax_results$fc, start = c(2022, 1), frequency = 12))
# ts plot of holdout set
plot(ts(toronto_holdout_ts, start = c(2022, 1), frequency = 12))

# 2. ACF PLOT
# ACF and PACF of training set
acf(toronto_train_ts, main = "ACF Plot: of Log-differenced NHPI: Toronto Training Set")
pacf(toronto_train_ts, main = "PACF Plot: of Log-differenced NHPI: Toronto Training Set")
# ACF from best ARIMA model
resid_fit <- residuals(toronto_transformed_nphi_auto)
acf(resid_fit, main = "ACF of ARIMA(1,0,2)(1,0,0)[12] residuals")
# ACF from best ARIMAX model
resid_fit <- residuals(toronto_fit_arimax)
acf(resid_fit, main = "ACF of ARIMAX(1,0,2)(1,0,0)[12] residuals")

# 3. SUMMARY STATISTICS
summary(toronto_train_ts)
summary(toronto_holdout_ts)
# summary of best ARIMA model
summary(toronto_transformed_nphi_auto)
# summary of best ARIMAX model
summary(toronto_fit_arimax)
# ccf between log-differenced NHPI and the lagged interest rate
ccf(toronto_irlag1[-1], diff(log(toronto_nhpi)))

# 4. RMSE
rmse = cbind(toronto_persist_fc$rmse, toronto_iid_fc$rmse,
             toronto_esm_fc$rmse, arima_results$rmse, arimax_results$rmse)
colnames(rmse) = c("persist", "iid", "esm", "arima", "arimax")
print(round(rmse, 3))
```


## Analyzing Vancouver

### Setting up
```{r}
# Reading the data
vancouver_nhpi_data <- read.csv("Full_Cities/vancouver_full.csv")
vancouver_nhpi_ir_data <- read.csv("Extracted_Cities/vancouver.csv")

# Extracting the NHPI
vancouver_nhpi <- vancouver_nhpi_data$VALUE

# Extracting the one-month lag interest rate
vancouver_irlag1 <- vancouver_nhpi_ir_data$IR_lag1

# Setting the time window from 1981-2023
vancouver_ts <- ts(vancouver_nhpi, start = c(1981,1), end = c(2023,12), frequency = 12)
vancouver_ir_ts <- ts(vancouver_irlag1, start = c(1981,1), end = c(2023,12), frequency = 12)

# Applying a transformation for NHPI
vancouver_transformed_nhpi <- diff(log(vancouver_ts))

# Setting a training set from 1981-2021
vancouver_train_ts <- window(vancouver_transformed_nhpi, end = c(2021,12))
vancouver_train_ir_xreg <- window(vancouver_ir_ts, start = c(1981,2), end = c(2021,12))

# Setting a holdout set from 2022-2023
vancouver_holdout_ts <- window(vancouver_transformed_nhpi, start = c(2022,1))
vancouver_holdout_ir_xreg <- window(vancouver_ir_ts, start = c(2022,1))
```

### Doing the baseline forecasts
```{r}
# Baseline 1: Naive/Persistence
vancouver_persist_fc <- persist_fc(vancouver_train_ts,vancouver_holdout_ts, iprint = FALSE)

# Baseline 2: Training-Set Average
vancouver_iid_fc <- iid_fc(vancouver_train_ts, vancouver_holdout_ts, iprint = FALSE)
```

### Doing the exponential smoothing forecasts
```{r}
library(forecast)

# Method 1
ntrain <- length(vancouver_train_ts)
ntotal <- length(vancouver_transformed_nhpi)

vancouver_transformed_nhpi_train <- vancouver_transformed_nhpi[1:ntrain]
vancouver_transformed_nhpi_holdout <- vancouver_transformed_nhpi[(ntrain+1):ntotal]

# Simple Exponential Smoothing
fit_expsmo <- HoltWinters(vancouver_train_ts, beta = FALSE, gamma = FALSE)
vancouver_esm_fc <- esm_fc(vancouver_transformed_nhpi_train,
                           vancouver_transformed_nhpi_holdout,
                           alpha = fit_expsmo$alpha,
                           level = fit_expsmo$coefficients["a"],
                           iprint = FALSE)

# Linear Exponential Smoothing
fit_hw <- HoltWinters(vancouver_train_ts, gamma = FALSE)
vancouver_holt_fc <- lholt_fc(vancouver_transformed_nhpi_train,
                              vancouver_transformed_nhpi_holdout,
                              alpha = fit_hw$alpha,
                              beta = fit_hw$beta,
                              level = fit_hw$coefficients["a"],
                              slope = fit_hw$coefficients["b"],
                              iprint = FALSE)

# Additive Seasonal Exponential Smoothing (using ETS for stability)
# We switched to ets() instead of HoltWinters() because it failed to converage for the Vancouver series.
vancouver_aseason_fit <- ets(vancouver_train_ts, model = "AAA")
summary(vancouver_aseason_fit)

vancouver_aseason_forecast <- forecast(vancouver_aseason_fit,
                                       h = length(vancouver_holdout_ts))

vancouver_aseason_rmse <- sqrt(mean((vancouver_holdout_ts -
                                     vancouver_aseason_forecast$mean)^2))

# Comparing RMSE
rmse_esm_vancouver <- cbind(vancouver_esm_fc$rmse,
                            vancouver_holt_fc$rmse,
                            vancouver_aseason_rmse)
colnames(rmse_esm_vancouver) <- c("esm", "lholt", "aseason")
rmse_esm_vancouver

# Best exponential smoothing method: Simple Exponential Smoothing
```

### Doing the ARIMA forecasts
```{r}
# Method 2
vancouver_auto_arima <- forecast::auto.arima(vancouver_train_ts)
summary(vancouver_auto_arima)

# Check residual ACF
acf(residuals(vancouver_auto_arima),
    main = "ACF of Vancouver auto.arima residuals")

# Fit several nearby candidate ARIMA models manually

fit1 <- arima(vancouver_train_ts,
              order = c(1,0,1),
              seasonal = list(order = c(1,0,0), period = 12),
              method = "ML")

fit2 <- arima(vancouver_train_ts,
              order = c(1,0,2),
              seasonal = list(order = c(1,0,0), period = 12),
              method = "ML")

fit3 <- arima(vancouver_train_ts,
              order = c(2,0,1),
              seasonal = list(order = c(1,0,0), period = 12),
              method = "ML")

fit4 <- arima(vancouver_train_ts,
              order = c(2,0,2),
              seasonal = list(order = c(1,0,0), period = 12),
              method = "ML")

fit5 <- arima(vancouver_train_ts,
              order = c(1,0,2),
              seasonal = list(order = c(1,0,1), period = 12),
              method = "ML")

# Compare AIC values
aic_vancouver <- cbind(fit1$aic, fit2$aic, fit3$aic, fit4$aic, fit5$aic)
colnames(aic_vancouver) <- c("fit1", "fit2", "fit3", "fit4", "fit5")
aic_vancouver

# Check residual ACFs
acf(residuals(fit1), main = "ACF of fit1 residuals")
acf(residuals(fit2), main = "ACF of fit2 residuals")
acf(residuals(fit3), main = "ACF of fit3 residuals")
acf(residuals(fit4), main = "ACF of fit4 residuals")
acf(residuals(fit5), main = "ACF of fit5 residuals")

# Choose the best candidate based on lowest AIC and ACF plot
# fit3 has smallest aic, so we choose fit3
order <- c(2,0,1)
seasonal <- list(order = c(1,0,0), period = 12)

# Fit the selected model on the training set
train_arima_model <- arima(vancouver_train_ts,
                           order = order,
                           seasonal = seasonal,
                           method = "ML",
                           include.mean = TRUE)

# Moving 1-step ahead forecasts for the holdout set
vancouver_arima_results <- arma_fc(tsdata = vancouver_transformed_nhpi,
                                   ntrain = length(vancouver_train_ts),
                                   order = order,
                                   seasonal = seasonal,
                                   method = "ML",
                                   traincoef = coef(train_arima_model),
                                   include.mean = TRUE,
                                   iprint = FALSE)

vancouver_arima_results$rmse
```
### Compare models using holdout rmse
```{r}
# Compare candidate ARIMA models using holdout RMSE

rmse_fit1 <- arma_fc(
  tsdata = vancouver_transformed_nhpi,
  ntrain = length(vancouver_train_ts),
  order = c(1,0,1),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(fit1),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_fit2 <- arma_fc(
  tsdata = vancouver_transformed_nhpi,
  ntrain = length(vancouver_train_ts),
  order = c(1,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(fit2),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_fit3 <- arma_fc(
  tsdata = vancouver_transformed_nhpi,
  ntrain = length(vancouver_train_ts),
  order = c(2,0,1),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(fit3),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_fit4 <- arma_fc(
  tsdata = vancouver_transformed_nhpi,
  ntrain = length(vancouver_train_ts),
  order = c(2,0,2),
  seasonal = list(order = c(1,0,0), period = 12),
  method = "ML",
  traincoef = coef(fit4),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

rmse_fit5 <- arma_fc(
  tsdata = vancouver_transformed_nhpi,
  ntrain = length(vancouver_train_ts),
  order = c(1,0,2),
  seasonal = list(order = c(1,0,1), period = 12),
  method = "ML",
  traincoef = coef(fit5),
  include.mean = TRUE,
  iprint = FALSE
)$rmse

# Put all candidate RMSE values into one table
arima_rmse_compare <- data.frame(
  model = c(
    "ARIMA(1,0,1)(1,0,0)[12]",
    "ARIMA(1,0,2)(1,0,0)[12]",
    "ARIMA(2,0,1)(1,0,0)[12]",
    "ARIMA(2,0,2)(1,0,0)[12]",
    "ARIMA(1,0,2)(1,0,1)[12]"
  ),
  rmse = c(
    rmse_fit1,
    rmse_fit2,
    rmse_fit3,
    rmse_fit4,
    rmse_fit5
  )
)

arima_rmse_compare

best_arima_model_V <- "ARIMA(1,0,2)(1,0,1)[12]"
best_arima_rmse_V <- 0.004501499

```


### Doing the ARIMAX forecasts with lagged interest
```{r}
# Method 3
vancouver_fit_arimax <- forecast::auto.arima(vancouver_train_ts,
                                             xreg = vancouver_train_ir_xreg)
summary(vancouver_fit_arimax)

acf(residuals(vancouver_fit_arimax),
    main = "ACF of Vancouver ARIMAX residuals")

order <- c(1,0,2)
seasonal <- list(order = c(1,0,1), period = 12)

# Use the same ARIMA order as chosen above as a starting point
train_arimax_model_vancouver <- arima(vancouver_train_ts,
                                      order = order,
                                      seasonal = seasonal,
                                      xreg = vancouver_train_ir_xreg,
                                      method = "ML",
                                      include.mean = TRUE)

# Getting the result of the moving 1-step ahead forecast for holdout set
vancouver_arimax_results <- armax_fc(tsdata = vancouver_transformed_nhpi, ntrain = length(vancouver_train_ts),
                                     order = order,
                                     seasonal = seasonal,
                                     xreg = vancouver_irlag1[-1],
                                     method = "ML",
                                     traincoef = coef(train_arimax_model_vancouver), include.mean = TRUE)

vancouver_arimax_results$rmse

best_arimax_model_V <- "ARIMAX(1,0,2)(1,0,1)[12]" 
best_arimax_rmse_V  <- vancouver_arimax_results$rmse

vancouver_best <- data.frame(
  City = "Vancouver",
  ARIMA_Model = best_arima_model_V,
  ARIMA_RMSE = round(best_arima_rmse_V, 3),
  ARIMAX_Model = best_arimax_model_V,
  ARIMAX_RMSE = round(best_arimax_rmse_V, 3)
)
```

###  Viewing the results of Vancouver's analyses
```{r}
# 1. TIME SERIES PLOT
plot(vancouver_ts, main = "Vancouver NHPI (Level)")
plot(vancouver_transformed_nhpi, main = "Vancouver NHPI Log-Difference")

# Forecast plots
plot(ts(vancouver_persist_fc$fc, start = c(2022,1), frequency = 12),
     main = "Vancouver Naive Forecast")
plot(ts(vancouver_iid_fc$fc, start = c(2022,1), frequency = 12),
     main = "Vancouver Training Mean Forecast")
plot(ts(vancouver_esm_fc$fc, start = c(2022,1), frequency = 12),
     main = "Vancouver Exponential Smoothing Forecast")
plot(ts(vancouver_arima_results$fc, start = c(2022,1), frequency = 12),
     main = "Vancouver ARIMA Forecast")
plot(ts(vancouver_arimax_results$fc, start = c(2022,1), frequency = 12),
     main = "Vancouver ARIMAX Forecast")
plot(ts(vancouver_holdout_ts, start = c(2022,1), frequency = 12),
     main = "Vancouver Holdout Series")

# 2. ACF PLOT
acf(vancouver_train_ts,
    main = "ACF Plot: Log-differenced NHPI: Vancouver Training Set")
pacf(vancouver_train_ts,
     main = "PACF Plot: Log-differenced NHPI: Vancouver Training Set")

acf(residuals(vancouver_auto_arima),
    main = "ACF of ARIMA(1,0,2)(1,0,1)[12] residuals")

acf(residuals(vancouver_fit_arimax),
    main = "ACF of ARIMA(1,0,2)(1,0,1)[12] residuals")

# 3. SUMMARY STATISTICS
summary(vancouver_train_ts)
summary(vancouver_holdout_ts)
summary(vancouver_auto_arima)
summary(vancouver_fit_arimax)

# CCF between log-differenced NHPI and lagged interest rate
ccf(vancouver_irlag1[-1], diff(log(vancouver_nhpi)),
    main = "CCF: Lagged Interest Rate vs Log-Differenced Vancouver NHPI")

# 4. RMSE
vancouver_rmse <- cbind(vancouver_persist_fc$rmse,
                        vancouver_iid_fc$rmse,
                        vancouver_esm_fc$rmse,
                        vancouver_arima_results$rmse,
                        vancouver_arimax_results$rmse)
colnames(vancouver_rmse) <- c("persist", "iid", "esm", "arima", "arimax")
print(round(vancouver_rmse, 3))
```
### Create the best ARIMA and ARMAX/ARIMAX models that found based on holdout set RMSE
```{r}
best_model_table <- rbind(
  montreal_best,
  toronto_best,
  vancouver_best
)

best_model_table

```

## Final Comparison Table
```{r}
library(knitr)

results_table <- data.frame(
  City   = c("Montreal", "Toronto", "Vancouver"),
  Persist = c(montreal_persist_fc$rmse,
              toronto_persist_fc$rmse,
              vancouver_persist_fc$rmse),
  IID    = c(montreal_iid_fc$rmse,
             toronto_iid_fc$rmse,
             vancouver_iid_fc$rmse),
  ESM    = c(montreal_holt_fc$rmse,    
             toronto_esm_fc$rmse,      
             vancouver_esm_fc$rmse),    
  ARIMA  = c(montreal_arima_results$rmse,
             toronto_arima_results$rmse,
             vancouver_arima_results$rmse),
  ARIMAX = c(montreal_arimax_results$rmse,
             toronto_arimax_results$rmse,
             vancouver_arimax_results$rmse)
)

# Average row
avg_row <- data.frame(
  City   = "Average",
  Persist = mean(results_table$Persist),
  IID    = mean(results_table$IID),
  ESM    = mean(results_table$ESM),
  ARIMA  = mean(results_table$ARIMA),
  ARIMAX = mean(results_table$ARIMAX)
)

final_table <- rbind(results_table, avg_row)

kable(final_table, digits = 3,
      caption = "One-Step-Ahead Rolling Forecast RMSE by City and Method")

# Best overall method
avg_rmse <- colMeans(results_table[, -1])
best_method <- names(which.min(avg_rmse))
cat("Best overall method:", best_method,
    "| Average RMSE =", round(min(avg_rmse), 3), "\n")

```

